Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-09-14 write_up/mixfishsim_ecolmod.tex
revising abstract

Most fisheries exploit a variety of spatially and temporally heterogeneous fish
populations using species-unselective gear. This can result in unintended,
unwanted catch of low quota or protected species and reducing these catches is
crucial for biological and economic sustainability of `mixed fisheries'. \\
%%
If fisheries are to avoid unwanted catch, an in-depth understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions and movement, and how fishers interact with different fish
populations. Because data on fish location at high temporal and spatial
resolutions is expensive and difficult to collect proxies are usually inferred
from scientific surveys or commercial catches with limited spatial and temporal
resolution. \\ 
%%
To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal
event-based simulation model incorporating: i) delay-difference
population dynamics, ii) population movement using Gaussian Random Fields to
simulate patchy, heterogeneously distributed populations, and iii) fishery
dynamics for multiple fleet characteristics based on species targeting under an
explore-exploit strategy. This is implemented via a mix of correlated random
walk movement (for exploration) and learned behaviour (for exploitation) phases
of the fisheries.  \\ 
%%
Fifty years of sub-daily fishing was simulated and used to draw inference on
the underlying community structures. We compared inferences based on:
commercial catch, a simulated fixed-site sampling survey design and the true
underlying community. We i) establish the potential limitations of
fishery-dependent data in providing a robust picture of spatiotemporal
distributions; and then ii) simulated an area closure based on areas defined
from the different data sources at a range of temporal and spatial resolutions.
\\
%%
Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our example framework
application that commercial data, while containing bias, provide a useful tool
for managing catches in mixed fisheries if applied at the correct
spatiotemporal scale. [NEED A LITTLE MORE DETAIL OF THE RESULTS HERE - 2 MORE
SENTENCES ON KEY TAKE-HOME RESULTS]\\
%%
\end{abstract}

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions.
In doing so fisheries catch an assemblage of species and may discard over-quota
catch when managed by single species quotas and fishers exhaust one or more
quota. This may lead to overexploitation of fish populations
\citep{Ulrich2011a, Batsleer2015}. Discarding of fish in excess of quota limits
the ability to maintain fishing mortality within sustainable limits
\citep{Alverson1994, Crowder1998, Rijnsdorp2007} and the ability to manage for
the biological and economic sustainability of fisheries. As such, there is
increasing interest in technical solutions such as gear and spatial closures as
measures to reduce unwanted catch \citep{Kennelly2002, Catchpole2008,
	Bellido2011, Cosgrove2019}. \\

Adaptive spatial management strategies have been proposed as a way of reducing
over-quota discards \citep{Holmes2011, Little2014, Dunn2014a}. However,
implementation is restricted by lack of knowledge of fish and fishery
spatiotemporal dynamics and understanding of the scale at which processes
become important for management. Understanding the correct scale for spatial
measures is crucial for implementing effective solutions which  minimising
economic impact \citep{Dunn2016}. For example, the problem can be to identify a
scale that promotes species avoidance for vulnerable or low quota species while
allowing continuance of sustainable fisheries for available quota species.\\

Identifying appropriate spatial scales for fisheries closures has been a
challenge in the past but is crucial to their success \citep{Costello2010,
	Dunn2016}. Inference is hampered where spatial information is coarse
due to low resolution reporting of fisheries catch which is aggregated across
larger gridded areas \citep{Branch2005}. Further, if data does not allow
identification of spatial features it may lead to poorly sited closures which
are ineffectual or have unintended consequences. For example, increased benthic
impact on previously unexploited areas from the cod closure in the North Sea
were observed without the intended effect of reducing cod exploitation
\citep{Rijnsdorp2001,Dinmore2003}). More refined spatiotemporal information has
since become available through the combination of logbook and Vessel Monitoring
System (VMS) data \citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and
more real-time spatial management has been possible \citep[e.g.][]{Holmes2011}.
However, fishers establish favoured fishing grounds through an explore-exploit
strategy \citep{Bailey2018} where they search for areas with high catches and
then use experience to return to areas where they've experienced high catch in
the past.  This leads to an inherently biased sampling where target species are
over-represented in the catch as fishers exploit areas of high abundance. \\

>>>>

%%
To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal
event-based simulation model incorporating: i) delay-difference
population dynamics, ii) population movement using Gaussian Random Fields to
simulate patchy, heterogeneously distributed populations, and iii) fishery
dynamics for multiple fleet characteristics based on species targeting under an
explore-exploit strategy. This is implemented via a mix of correlated random
walk movement (for exploration) and learned behaviour (for exploitation) phases
of the fisheries.  \\ 
%%
Fifty years of sub-daily fishing was simulated and used to draw inference on
the underlying community structures. We compared inferences based on:
commercial catch, a simulated fixed-site sampling survey design and the true
underlying community. We i) establish the potential limitations of
fishery-dependent data in providing a robust picture of spatiotemporal
distributions; and then ii) simulated an area closure based on areas defined
from the different data sources at a range of temporal and spatial resolutions.
\\
%%
Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our example framework
application that commercial data, while containing bias, provide a useful tool
for managing catches in mixed fisheries if applied at the correct
spatiotemporal scale. [NEED A LITTLE MORE DETAIL OF THE RESULTS HERE - 2 MORE
SENTENCES ON KEY TAKE-HOME RESULTS]\\
%%
\end{abstract}

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time with varying knowledge of species distributions.
As fishers do not have full control over what species they select in `mixed
fisheries' this can result in catch of low quota or protected species. This
leads to discarding of over-quota catch if it is a species for which they have
no quota, limiting control over fishing mortality \citep{Alverson1994,
	Crowder1998, Rijnsdorp2007} and the ability to manage fisheries for the
biological and economic sustainability \citep{Ulrich2011a, Batsleer2015}.\\

There is increasing interest in technical solutions such as gear and spatial
closures as measures to reduce unwanted catch \citep{Kennelly2002,
	Catchpole2008, Bellido2011, Cosgrove2019} and adaptive spatial
management strategies have been proposed as a way of reducing over-quota
discards \citep{Holmes2011, Little2014, Dunn2014a}. However, if fisheries are
to reduce unwanted catch through spatial avoidance, an in-depth understanding
of spatiotemporal fishery dynamics is required. Implementation of spatial
measures is hampered by a lack of knowledge of fish and fishery spatiotemporal
dynamics and understanding of the scale at which processes become important for
management. \\

Understanding the correct scale for spatial measures is also crucial as it
enables implementation of effective solutions which minimise economic impact
\citep{Dunn2016}. For example, the problem can be to identify a scale that
promotes species avoidance for vulnerable or low quota species while allowing
continuance of sustainable fisheries for available quota species. Identifying
the correct spatial scale remains a challenge because data on fish location at
high temporal and spatial resolutions is expensive and difficult to collect and
proxies are usually inferred from scientific surveys or commercial catches with
limited spatial and temporal resolution. \\

Identifying appropriate spatial scales for fisheries closures has been a
highlighted as crucial to their success \citep{Costello2010, Dunn2016}.
Inference is hampered where spatial information is coarse due to low resolution
reporting of fisheries catch which is aggregated across larger gridded areas
\citep{Branch2005}. Further, if data does not allow identification of spatial
features it may lead to poorly sited closures which are ineffectual or have
unintended consequences. For example, increased benthic impact on previously
unexploited areas from the cod closure in the North Sea were observed without
the intended effect of reducing cod exploitation
\citep{Rijnsdorp2001,Dinmore2003}).\\

More refined spatiotemporal information has since become available through the
combination of logbook and Vessel Monitoring System (VMS) data \citep{Lee2010,
	Bastardie2010, Gerritsen2012, Mateo2016} and more real-time spatial
management has been possible \citep[e.g.][]{Holmes2011}.  However, fishers
establish favoured fishing grounds through an explore-exploit strategy
\citep{Bailey2018} where they search for areas with high catches and then use
experience to return to areas where they've experienced high catch in the past.
This leads to an inherently biased sampling where target species are
over-represented in the catch as fishers exploit areas of high abundance. \\
